---
title: "ESDA - Chronic diseases"
author: "Anaïs Ladoy"
date: '2023-04-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/mnt/data/GEOSAN/RESEARCH PROJECTS/GEOCHRONIC @ LASIG (EPFL)/GEOSAN-geochronic/src")
```

```{r, include=FALSE}
library(tidyverse)
library(sf)
source('/mnt/data/GEOSAN/FUNCTIONS/GIRAPH-functions/geosan_funcs/password_utils.R')
source('descriptive/utils_spatial_clustering_outcomes.R')
```


```{r, include=FALSE}
con <- dbConnect(drv=RPostgreSQL::PostgreSQL(),host = "localhost",user= "aladoy",rstudioapi::askForPassword(),dbname="geosan")
```


**Import the datasets**

- `indiv`: Subset of Colaus participants eligible for this study (see selection procedure) within Lausanne.
- `laus$extent`: Boundary of the Lausanne area (polygon).
- `laus$sectors`: Statistical sectors of Lausanne (for mapping purpose only).
- `env`: Living environment (hectares) characteristics.

```{r}
indiv <- load_participants(con)
laus <- load_boundaries(con)
env <- load_env()
```


***Specify covariates***

```{r}
cov.indiv <- c("age", "sex", "swiss", "cohabiting", "education", "difficulties", "smoking", "drinking", "inactivity")

cov.env <- c("PTOT", "ENV_INDEX", "NO2", "PM25", "NOISE", "GREEN_SP", "D_SPORT", "N_ACC_PED", "SOC_ECO_INDEX", "UNEMPLOYMENT", "LOW_EDUC", "HH_1PERS", "INCOME")
```



# Diabetes

```{r}
diab.data <- select_outcome(indiv, "diabetes", cov=cov.indiv)
```

Visualize diabetes events

```{r, echo=FALSE}
plot_case_control(diab.data, window=laus$extent, title='Diabetes')
```

```{r, include=FALSE}
diab.events <- create_ppp(diab.data, laus$extent, marks=diab.data$outcome, title="Diabetes - events")
diab.cases <- create_ppp(diab.data %>% filter(outcome=='case'), laus$extent, title="Diabetes - cases")
diab.controls <- create_ppp(diab.data %>% filter(outcome=='control'), laus$extent, title="Diabetes - controls")
```

**Explore the spatial density of diabetes.**
```{r}
compare_spatial_dens(diab.cases, diab.controls)
```

To construct the spatial density of diabetes, we used a two-dimensional product kernels using a Gaussian function and a bandwith defined by the Scott's rule (*see p.133-136, Waller and Gotway (2004) for more details*). We applied an edge correction using the Jones-Diggle formula.
The density function estimates suggest some local differences in the case and control patterns, especially near Montelly & Pontaise (high density of cases) and Chailly (high density of controls).

**Log ratio of spatial densities.**  

Detect local clusters (*see p.164-171, Waller and Gotway (2004) for more details*).
"Do cases and controls tend to occur in the same locations?"
Assumption: bot case and control locations each represent a realization of a heterogeneous Poisson point process over the study area.
We obtain a map indicating areas where events appear more or less likely to be cases than controls.
Bandwith: We choosed a bandwith of 350m which is a compromise between the Scott's rule estimated bandwith for cases and controls separatly.

```{r}
diab.lrr <- log_ratio_spatial_dens(diab.events, sigma=350)
```

In the last map, orange areas represent parts of the study area where cases are clustered relative to controls beyond what is expected according to the random labeling hypothesis (tolerance=0.9, nsim=999).
For the diabetes in Lausanne, we confirm our suggestions that there are clusters of cases near Montelly and Pontaise, and a deficit of cases near Chailly.
However, there is no evidence of global clustering (p=0.425).

**Difference between K-functions of case and control processes.**  

Using K-functions, we are interested of the distance to which any observed clustering tends to occur, averaged over the entire study area.
"Do cases tend to occur near other cases in the same manner that controls tend to occur near other controls?"

```{r}
k_func_diff(diab.events)
```


### Extract areas of significant higher and lower risks of diabetes

Polygonize the image of the log relative risk surface.
```{r, include=FALSE}
diab.incid <- polygonize_logrr(diab.lrr, env)
```

```{r}
map_significant_areas(diab.incid$lrr.poly, laus$sectors)
```


Compare neighborhood characteristics between areas of reduced and higher risk of diabetes.
```{r}
compare_risk_areas(diab.incid$lrr.ha, cov.env)
```

```{r}
compare_risk_groups(diab.incid$lrr.ha, cov.env)
```

**Summary**

While there is no evidence of global clustering of diabetes in Lausanne, we detected areas of significantly higher and lower risk.  

*Areas of higher risk:*  

- Montoie / Provence / Sébeillon
- Bossons / Bois-Mermet
- Bourdonnette

*Areas of lower risk:*    

- Bellerive
- Victor-Ruffy / Sallaz / Devin / Craivavers / Av. Secrétan

*Areas of higher risk are characterized by:*    

- higher population density (p<0.001)
- higher environmental exposures such as road noise and air pollution (p<0.001)
- lower greenspace (p<0.001)
- increased access to sport facility (p<0.001)
- lower neighborhood safety - higher pedestrian accidents (p<0.001) 
- higher socio-economic vulnerability (p=0.001)
- higher unemployment rate (p<0.001)
- higher rate of low education (p<0.001)
- lower income (p<0.001)


# Morbidity


```{r}
morbi.data <- select_outcome(indiv, "morbidity", cov=cov.indiv)
```

```{r, echo=FALSE}
plot_case_control(morbi.data, window=laus$extent, title='Morbidity')
```

```{r, include=FALSE}
morbi.events <- create_ppp(morbi.data, laus$extent, marks=morbi.data$outcome, title="Morbidity - events")
morbi.cases <- create_ppp(morbi.data %>% filter(outcome=='case'), laus$extent, title="Morbidity - cases")
morbi.controls <- create_ppp(morbi.data %>% filter(outcome=='control'), laus$extent, title="Morbidity - controls")
```

```{r}
compare_spatial_dens(morbi.cases, morbi.controls)
```
```{r}
morbi.lrr <- log_ratio_spatial_dens(morbi.events, sigma=300)
```

```{r}
k_func_diff(morbi.events)
```

```{r}
morbi.incid <- polygonize_logrr(morbi.lrr, env)
map_significant_areas(morbi.incid$lrr.poly, laus$sectors)
```

```{r}
compare_risk_areas(morbi.incid$lrr.ha, cov.env)
```
```{r}
compare_risk_groups(morbi.incid$lrr.ha, cov.env)
```

**Summary**

While there is no evidence of global clustering of diabetes in Lausanne, we detected areas of significantly higher and lower risk.  

*Areas of higher risk:*  

- Malley / Montoie / Provence / Sébeillon
- Bossons / Bois-Mermet / Ancien-Stand / Bellevaux/ Sauvabelin
- Chailly / Victor-Ruffy

*Areas of lower risk:*    

- Mont-Repos / Béthusy / Av. Sécrétan
- Riponne / Chauderon / Flon / Montbenon / Tivoli / Marc-Dufour / Cour / Milan / Bellerive

*Areas of higher risk are characterized by:*    

- higher population density (p<0.001)
- higher environmental exposures (p<0.001) but if we look at the exposures separately, we observe significant differences in air pollutants - NO2 (p<0.001) and PM2.5 (p<0.001) but no differences in road noise (p=0.3).
- higher greenspace (p<0.001)
- lower access to sport facility (p<0.001)
- higher neighborhood safety - lower pedestrian accidents (p<0.001) 
- higher socio-economic vulnerability (p<0.001)
- higher unemployment rate (p<0.001)
- higher rate of low education (p<0.001)
- lower income (p<0.001)

Compared to diabetes, areas of higher risk of morbidity (all causes) do not cumulate socio-economic and environmental exposures. Lower risk areas correspond to highly urbanized neighborhood (higher pedestrian accidents, lower greenspace, higher air pollution) but are still of lower population density and lower socio-economic vulnerability than high-risk areas.


# Multimorbidity

```{r}
multimorbi.data <- select_outcome(indiv, "multimorbidity", cov=cov.indiv)
```

```{r}
plot_case_control(multimorbi.data, window=laus$extent, title='Multimorbidity')
```


```{r, include=FALSE}
multimorbi.events <- create_ppp(multimorbi.data, laus$extent, marks=multimorbi.data$outcome, title="Multimorbidity - events")
multimorbi.cases <- create_ppp(multimorbi.data %>% filter(outcome=='case'), laus$extent, title="Multimorbidity - cases")
multimorbi.controls <- create_ppp(multimorbi.data %>% filter(outcome=='control'), laus$extent, title="Multimorbidity - controls")
```

```{r}
compare_spatial_dens(multimorbi.cases, multimorbi.controls)
```
```{r}
multimorbi.lrr <- log_ratio_spatial_dens(multimorbi.events, sigma=300)
```

```{r}
k_func_diff(multimorbi.events)
```

```{r}
multimorbi.incid <- polygonize_logrr(multimorbi.lrr, env)
map_significant_areas(multimorbi.incid$lrr.poly, laus$sectors)
```

```{r}
compare_risk_areas(multimorbi.incid$lrr.ha, cov.env)
```
```{r}
compare_risk_groups(multimorbi.incid$lrr.ha, cov.env)
```

