---
title: "Modeling - Chronic diseases"
author: "Anaïs Ladoy"
date: '2023-04-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/mnt/data/GEOSAN/RESEARCH PROJECTS/GEOCHRONIC @ LASIG (EPFL)/GEOSAN-geochronic/src")
```

```{r, include=FALSE}
library(stats)
library(tmap)
library(leaflet)
library(lme4)
require(RPostgreSQL)
library(MASS)
library(MapGAM)
library(pROC)
library(corrplot)
library(mgcv)
library(gam)
library(caret)
library(jtools)
library(PrevMap)
source('/mnt/data/GEOSAN/FUNCTIONS/GIRAPH-functions/geosan_funcs/password_utils.R')
source('descriptive/esda_utils.R')
```

```{r, include=FALSE}
con <- dbConnect(drv=RPostgreSQL::PostgreSQL(),host = "localhost",user= "aladoy",rstudioapi::askForPassword(),dbname="geosan")
```

### Import the datasets

- `data`: All F2 Colaus participants eligible for this study (see selection procedure).

```{r, include=FALSE}
data <- load_participants(con)
```

- `boundary.laus`: Boundary of the Lausanne area (polygon)
- `boundary.vd`: Boundary of the Vaud state (polygon)

```{r, include=FALSE}
extent <- load_boundaries(con)
```

- `env`: Neighborhood characteristics build in **explore_neighborhood_characteristics.Rmd**

```{r, include=FALSE}
env.sf <- read_sf("../processed_data/neighborhood_vd.gpkg")
```

- `predict.grid`: Prediction grid for Vaud state (inhabited hectares only)

```{r, include=FALSE}
grid.sf <- read_sf(con, query="SELECT * FROM vd_reli_centroid")
```


### Data Wrangling


<!-- Create a measure of neighborhood environmental exposure and deprivation by converting ENV_INDEX and SOC_ECO_INDEX to quintiles (n=5). -->
<!-- ```{r, include=FALSE} -->
<!-- env.sf <- env.sf %>% mutate(env_depriv = factor(ntile(env_index, 5)), soceco_depriv = factor(ntile(soc_eco_index, 5)), ptot_quint = factor(ntile(ptot, 5))) -->
<!-- ``` -->

Merge data with environmental covariates
```{r}
data <- left_join(data, env.sf %>% st_drop_geometry(), by="reli")
paste("Number of participants removed because no covariate data for corresponding RELI: ", data %>% filter(is.na(D_SPORT)) %>% nrow())
data <- data %>% filter(!is.na(D_SPORT))
```

<!-- ```{r} -->
<!-- tmap_mode("view") -->
<!-- tm_shape(env.sf) + -->
<!--   tm_polygons("env_depriv", palette = "YlOrRd", alpha=0.8, title="Quintiles") + -->
<!--   tm_layout(title="Neihgborhood environmental vulnerability in Lausanne") -->
<!-- tm_shape(env.sf) + -->
<!--   tm_polygons("soceco_depriv", palette = "YlOrRd", alpha=0.8, title="Quintiles") + -->
<!--   tm_layout(title="Neihgborhood socio-economic deprivation in Lausanne") -->
<!-- tm_shape(env.sf) + -->
<!--   tm_polygons("ptot_quint", palette = "YlOrRd", alpha=0.8, title="Quintiles") + -->
<!--   tm_layout(title="Neihgborhood population density in Lausanne") -->
<!-- ``` -->


```{r}
cov.indiv <- c("age", "sex", "swiss", "cohabiting", "education", "working", "difficulties", "smoking", "drinking", "inactivity", "moved")
# cov.env <- c("ptot", "d_sport", "d_swim", "n_acc_ped", "pm25", "noise")
cov.env <- c("PTOT", "D_SPORT", "N_ACC_PED", "GREEN_SP", "ENV_INDEX", "SOC_ECO_INDEX", "SWISS", "UNEMPLOYMENT", "LOW_EDUC", "INCOME", "HH_1PERS", "F_45_54", "F_55_64", "F_65_74", "F_75_MORE", "M_45_54", "M_55_64", "M_65_74", "M_75_MORE")
```

I removed HIGH_EDUC because highly correlated with LOW_EDUC in the model but the spatial pattern is different (see spatial_clustering_outcomes). Should I include both?
Same for poverty and income. Easier to predict with a binary variable but we loose some granularity and both are correlated. What should we keep in the model? For now, I kept income.

### Diabetes

Convert binary / categorical variables
```{r}
diab.data <- data %>% 
  mutate_at(cov.indiv[! cov.indiv == "age"], as.factor) %>% 
  mutate(POVERTY = factor(POVERTY)) %>%
  mutate(diabetes = factor(diabetes, levels = c("0", "1"))) %>%
  dplyr::select(pt, diabetes, all_of(cov.indiv), coordx, coordy, all_of(cov.env), reli)
```

Split dataset (train + test)
```{r}
diab.dt <- create_train_dataset(diab.data, frac=0.8, seed=NULL)
diab.train <- diab.dt$train
diab.test <- diab.dt$test
```

Distribution of age and neighborhood-related characteristics between cases and controls.
```{r}
diab.data.long <- diab.data %>% dplyr::select(c(diabetes, age, cov.env[! cov.env == "POVERTY"])) %>% gather(key = "Variable", value = "Value", -diabetes)

ggplot(diab.data.long, aes(x = Value, y = diabetes)) +
  geom_violin() +
  labs(x="", y="Diabetes") + 
  facet_wrap(~ Variable, scales = "free_x") 
```


```{r}
m.univ <- glm(diabetes ~ sex, data=diab.data, family='binomial')
print_model_summary(m.univ)
```
- `Deviance residuals`: should be centered around 0 and symmetrical. 
- `Coefficients`: give the log-odds estimates. Effect size.
- `Dispersion parameter`: since the variance is derived from the estimated mean (compared to the linear regression where we estimate both the mean and the variance from the data), it can be underestimated and we can adjust this by adjusting the dispersion parameter.
- `Null and Residual Deviance`: can be used to compare models, compute R^2 and an overall p-value.
- `AIC`: Akaike Information Criterion, which is here the residual deviance adjusted for the number of parameters in the model. Can be used to compare models.
- `Fisher Scoring Iterations`: how quickly the glm() function converged on the maximum likelihood estimates for the coefficients.

**Compare models:** Better model will have lower residual deviance and lower AIC.

```{r, fig.width=3,fig.height=2}
paste("The probability that a man has diabetes is", inverse_logit(m.univ$coefficients[[1]]))
paste("The probability that a woman has diabetes is", inverse_logit(m.univ$coefficients[[1]] + m.univ$coefficients[[2]]))
paste("Women are", round(1/exp(m.univ$coefficients[[2]]),2), "less likely to have diabetes than a man.")
mcfadden(m.univ)
draw_predicted_prob(m.univ, diab.data, "diabetes")
```

**Specification error**: we must particularly check here if our model has all relevant predictors, and if the linear combinations of them is sufficient.


Investigate individual-level covariates using a backward stepwise logistic regression.
```{r}
diab.indiv.dat <- diab.data %>% dplyr::select(diabetes, all_of(cov.indiv))
glm.indiv <- glm(diabetes ~ ., data=diab.indiv.dat, family=binomial()) %>% 
  stepAIC(direction = "backward")
print_model_summary(glm.indiv)
```

Test if we should model age as non-linear to diabetes (with cubic spline)
```{r}
gam.indiv <- mgcv::gam(diabetes ~ s(age, bs="cr") + sex + swiss + education + working + difficulties + smoking + inactivity, method = "GCV.Cp", data=diab.data, family="binomial")
summary(gam.indiv)
```
```{r}
compare_models(glm.indiv, gam.indiv)
```
Plot the cubic splines used to model the diabetes-age relationship
```{r, fig.width=3,fig.height=2}
plot(gam.indiv, cex.lab=0.8, cex.axis=0.8)
```

```{r}
gam.indiv.final <- mgcv::gam(diabetes ~ s(age, bs="cr") + sex  + education + difficulties  + smoking + inactivity, method = "GCV.Cp", data=diab.data, family="binomial")
summary(gam.indiv.final)
```
```{r}
compare_models(gam.indiv, gam.indiv.final)
```


There are no significant differences between the GAM with all individual covariates selected from the backward stepwise regression, including a cubic-spline relationship between age and diabetes (gam.indiv), and the model with a reduced number of covariates (removing the ones where p-value > 0.05). Therefore, we keep the reduced version (gam.indiv.reduced).
-> there is a significant difference actually. Should we then preserved the full model returned by the stepwise regression even if some covariates are not significant?


### Assess the impact of the environment at the individual level

With socio-eco / environment and demographic characteristics
```{r}
glm.env.all <- glm(diabetes ~ ., data=diab.data %>% dplyr::select(diabetes, all_of(cov.env)), family=binomial()) %>% 
  stepAIC(direction = "backward", trace=0)
print_model_summary(glm.env.all)
```


With socio-eco / environment and without demographic characteristics
```{r}
glm.env.nodemo <- update(glm.env.all, ~ . - F_45_54 - F_55_64 - F_65_74 - F_75_MORE - M_45_54 - M_55_64 - M_65_74 - M_75_MORE )
print_model_summary(glm.env.nodemo)
```

Significant environmental variables only
```{r}
glm.env.final <- glm(diabetes ~  SOC_ECO_INDEX + LOW_EDUC + INCOME , data=diab.data, family=binomial())
print_model_summary(glm.env.final)
```

```{r}
compare_models(glm.env.nodemo, glm.env.final)
```



### Incorporate spatial effects

Try an empty 2-levels regression model. 
Determine to what extent do the log-odds vary between clusters using the intraclass correlation coefficient (ICC). The ICC quantifies the degree of homogeneity of the outcome within clusters, and represents the proportion of the between-cluster variation in the total variation.
```{r}
multilevel.empty <- glmer(diabetes ~ (1 | reli), data=diab.data, family=binomial("logit"))
summary(multilevel.empty)
icc <- multilevel.empty@theta[1]^2/ (multilevel.empty@theta[1]^2 + (3.14159^2/3))
icc
```

<!-- ```{r} -->
<!-- ggplot(diab.data, aes(ptot, env_index)) +  -->
<!--   geom_point() + -->
<!--   geom_smooth() -->
<!-- ggplot(diab.data, aes(ptot, soc_eco_index)) +  -->
<!--   geom_point() + -->
<!--   geom_smooth() -->
<!-- ``` -->


Incorporate neighborhood-level variables without hierarchical modeling.
```{r}
gam.spat.complete <- update(gam.indiv, ~ . + SOC_ECO_INDEX + LOW_EDUC + INCOME)
summary(gam.spat.complete)
exp(coef(gam.spat.complete))
```

```{r}
gam.spat.final <- update(gam.spat.complete, ~ . - INCOME - LOW_EDUC)
summary(gam.spat.final)
exp(coef(gam.spat.final))
```


For the same individual-level covariates, individuals living in poorer SEP are 11 times more likely to have diabetes. -> the odds-ratio seem particularly huge (?).

```{r}
compare_models(gam.indiv.final, gam.spat.final)
```


Generalized Additive Model (GAM) with a two-dimensional LOESS smooth.
```{r}
gam.loess <- update(gam.spat.final, ~ . + lo(coordx, coordy, span=0.3))
summary(gam.loess)
```

```{r}
compare_models(gam.spat.final, gam.loess)
```
Adding a LOESS spatial smoother does not improve significantly the model. We therefore keep the previous model.

<!-- ```{r} -->
<!-- a <- data.frame(pred_prob = inverse_logit(predict.gam(gam.spat.final, diab.data, type='response')), observed=diab.data$diabetes) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- predgrid <- predgrid(diab.data[, c("coordx","coordy")], map=as(boundary.vd, "Spatial")) -->
<!-- m1 <- modgam(diabetes~ age + sex + education + difficulties + inactivity + SOC_ECO_INDEX + lo(coordx, coordy), data=diab.data, rgrid=predgrid, sp=0.3, type="spatial", family = "binomial", reference="none", permute=0) -->
<!-- summary(m1) -->
<!-- plot(m1, exp=TRUE, data, contours = "response") -->
<!-- ``` -->



```{r, fig.width=5,fig.height=3}
draw_predicted_prob(gam.spat.final, diab.data, "diabetes")
```



Assess model accuracy by comparing predicted outcome at test locations (20% of the dataset).
```{r}
pred.indiv <- predict(m.indiv, newdata=test, type="response")
confusionMatrix(factor(as.numeric(pred.indiv>0.5)), test$diabetes)
```

```{r}
roc_curve <- roc(test$diabetes, pred.indiv)
plot(roc_curve)
auc(roc_curve)
```



```{r}
effect_plot(m, pred = age, plot.points = TRUE,
            jitter = c(0.1, 0.05), point.alpha = 0.1) +
  ylab("Pr(SmokeNow = Yes)")
```


